---
title: "RNAseq_production_cell"
output: html_notebook
---


##### Load all the libraries needed for the analysis and set working directory
```{r}
source("https://bioconductor.org/biocLite.R")

biocLite("limma","gplots","edgeR","org.Hs.eg.db","RColorBrewer","ALL","GO.db","annotate","genefilter","GOstats","RColorBrewer","xtable","DBI")
library("limma")
library("gplots")
library("edgeR")
library("org.Hs.eg.db")
library("RColorBrewer")
library("ALL")
library("GO.db")
library("annotate")
library("genefilter")
library("GOstats")
library("RColorBrewer")
library("xtable")
library("DBI")
setwd("/Users/jguo/AudentesProjects/RNAseq-production-cell-YA1706202/subread_mapping")
getwd()

```

##### Load read_count file generated by subread featureCount on DNAnexus, create experimental design and DGElist object (a common data format recognized by most RNAseq data analysis packages)
## the count table could be downloaded at https://dl.dnanex.us/F/D/BjYzQFz2K2115G41BfB8V6jV50p244f8x1jvKGgy/1_S1_R1_001_hg38_analysisSet_subread_ff_count.txt).
## note that the header line of the count table has to be removed to conform to the required input format.

```{r}

data = read.delim("1_S1_R1_001_hg38_analysisSet_subread_ff_count.txt")
head(data)
str(data)
# subset columns with Gene_ID, Gene_len, and counts)
data_subset = data[,c(1,6:13)]
head(data_subset)
# set up experimental design
exp_design = read.delim("exp_design.txt")
exp_design
lev = c("prior","0h_post","24h_post","48h_post","72h_post", "96h_post","120h_post")
f = factor(exp_design$Treatment, levels=lev)
head(f)
design = model.matrix(~0+f)
colnames(design) = lev
# Generate DGE table
dge = DGEList(counts=data[,c(7:13)],genes=data[,c("Geneid","Length")], remove.zeros =TRUE, group = f)
dge
## Get log2 counts per million
logcounts = cpm(dge, log=TRUE)
## MDS plot
plotMDS(dge)

# Hierarchical clustering with heatmaps
## estimate the variance for each gene across all samples
var_genes <- apply(logcounts, 1, var)
head(var_genes)
## Get the names for the top 500 most variable genes across all samples
select_var_100 = names(sort(var_genes, decreasing=TRUE))[1:100]
select_var_500 <- names(sort(var_genes, decreasing=TRUE))[1:500]

## subset logcounts matrix
highly_variable_lcpm_100 = logcounts[select_var_100,]
highly_variable_lcpm_500 = logcounts[select_var_500,]

dim(highly_variable_lcpm)
##set colours
mypalette <- brewer.pal(11,"RdYlBu")
morecols <- colorRampPalette(mypalette)

## generate heat map
heatmap.2(highly_variable_lcpm_100,Colv=FALSE,col=rev(morecols(50)),trace="none", main="Top 100 most variable genes across samples",scale="row")

heatmap.2(highly_variable_lcpm_500,Colv=FALSE,col=rev(morecols(50)),trace="none", main="Top 500 most variable genes across samples",scale="row")

# Gene set enrichment

## use all genes that are expressed at least in one sample as universal gene set
all_expressed_gene_id = names(sort(var_genes, decreasing=TRUE))

## Run gene set enrichment analysis with conditional and uncoditional settings 

params_100 = new('GOHyperGParams', geneIds=select_var_100, universeGeneIds=all_expressed_gene_id, annotation='org.Hs.eg.db', ontology='BP', pvalueCutoff=0.01, conditional=FALSE, testDirection='over')
hgOver_100 = hyperGTest(params_100)
htmlReport(hgOver_100, file = "hgOver_100.html")

params_conditional_100 = new('GOHyperGParams', geneIds=select_var_100, universeGeneIds=all_expressed_gene_id, annotation='org.Hs.eg.db', ontology='BP', pvalueCutoff=0.01, conditional=TRUE, testDirection='over')
hgConOver_100 = hyperGTest(params_conditional_100)
htmlReport(hgConOver_100, file = "hgConOver_100.html")

params_500 = new('GOHyperGParams', geneIds=select_var_500, universeGeneIds=all_expressed_gene_id, annotation='org.Hs.eg.db', ontology='BP', pvalueCutoff=0.01, conditional=FALSE, testDirection='over')
hgOver_500 = hyperGTest(params_500)
htmlReport(hgOver_500, file = "hgOver_500.html")

params_conditional_500 = new('GOHyperGParams', geneIds=select_var_500, universeGeneIds=all_expressed_gene_id, annotation='org.Hs.eg.db', ontology='BP', pvalueCutoff=0.01, conditional=TRUE, testDirection='over')
hgConOver_500 = hyperGTest(params_conditional_500)
htmlReport(hgConOver_500, file = "hgConOver_500.html")

sessionInfo()
```
####
R version 3.3.2 (2016-10-31)
Platform: x86_64-apple-darwin11.4.2 (64-bit)
Running under: macOS Sierra 10.12.6

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets
[8] methods   base

other attached packages:
 [1] DBI_0.7              xtable_1.8-2         GOstats_2.40.0
 [4] graph_1.52.0         Category_2.40.0      Matrix_1.2-12
 [7] genefilter_1.56.0    annotate_1.52.1      XML_3.98-1.5
[10] GO.db_3.4.0          ALL_1.16.0           RColorBrewer_1.1-2
[13] org.Hs.eg.db_3.4.0   AnnotationDbi_1.36.2 IRanges_2.8.2
[16] S4Vectors_0.12.2     Biobase_2.34.0       BiocGenerics_0.20.0
[19] edgeR_3.16.5         gplots_3.0.1         limma_3.30.13

loaded via a namespace (and not attached):
 [1] Rcpp_0.12.14           bitops_1.0-6           digest_0.6.13
 [4] bit_1.1-12             RSQLite_2.0            memoise_1.1.0
 [7] tibble_1.3.4           lattice_0.20-35        pkgconfig_2.0.1
[10] rlang_0.1.4            gtools_3.5.0           caTools_1.17.1
[13] locfit_1.5-9.1         bit64_0.9-7            grid_3.3.2
[16] GSEABase_1.36.0        RBGL_1.50.0            survival_2.41-3
[19] gdata_2.18.0           blob_1.1.0             splines_3.3.2
[22] AnnotationForge_1.16.1 KernSmooth_2.23-15     RCurl_1.95-4.8
####

